---
title: "Photorespiration-Species"
author: "Philip"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Before using any of the code below, make sure that all of the necessary packages are installed using this line. If any packages are missing from the library, please do install them manually using the hashed out line below. 

```{r}
install.packages("Hmisc", dependencies = T) 

# install.packages("missingpackage")
```




#Loading in all the packages and setting up so that all variables are expressed correctly


```{r}

packages_to_load <- c("drc", "bbmle", "labelled", "MALDIquant", 
                      "magicfor", "ggplot2", "ggpubr", "ggpmisc", "tidyverse", "broom", 
                      "Hmisc", "plotly", "PairedData", "DescTools", 
                      "generalhoslem", "graphics", "nls2", "plyr",
                      "ggthemes", "ggpubr", "signal", "segmented")

lapply(packages_to_load, library, character.only = TRUE)

## lable ----
ytleaf          <- expression("T"["leaf"]~degree~"C")
photorate       <- expression("" ~mu ~ mol ~ "CO"[2]~m^{-2} ~ "s"^{-1})
labrsq          <- expression(R^{2}~"=")
labvpdl         <- "VPD (kPa)"
lab_photo       <- expression(paste(italic(A)[Net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')'))
lab_photo1      <- expression(paste(italic(A)[Net], ' (', mu * 'mol m'^{-2}*'s'^{-1}*')'))
lab_Tleaf       <- expression(paste("Leaf temperature (", degree,"C)"))
labcond         <- expression("g"["s"]~~"in"~mol~H[2]*O~m^{-2}~s^{-1})
labcond1        <- expression("Conductance to "~H[2]*O~"in"~mol~H[2]*O~m^{-2}~s^{-1})
labtrmmol       <- expression("E"~~"in"~mol~H[2]*O~m^{-2}~s^{-1})
etrrate         <- expression("" ~mu ~ mol ~"photon" ~m^{-2} ~ "s"^{-1})
photorate       <- expression("" ~mu ~ mol ~ "CO"[2]~m^{-2} ~ "s"^{-1})
labgs           <- expression(paste(italic(g[s])*" ("~mol[H[2]*O]~m^{-2}~s^{-1}~")"))



```


#Loading in the compiled species data into a dataframe (df).

```{r}
df <- read.csv("Data/Uppsala-2024-Summer-Photorespiration-SpotMes-TreeSpp.csv", header = T, stringsAsFactors = T, sep = ";")
```



#Cleaning up the dataframe by sorting out any entries in the loaded file that will not be used in the analysis. Thereafter pairing entries together (O2-free with ambient for the same replicate). Then converting all the pairs into factors.

```{r}
df     <- subset(df, dq == "yes") # eliminates unreliable values.

df$pairid <- paste0(df$sprp, "-", df$setTleaf)

df$pairid <- as.factor(df$pairid)


length(unique(df$pairid))       # individual runs + season

dflist    <-   split(df,list(df$pairid))
dflist    <-   split(df,unique(list(df$pairid)))
dflist    <-   dflist[sapply(dflist, nrow)>0] 

setwd("output")
```



#Next, running the for loop that will give us our worksheet for the statistical analysis. All variables that we want to calculate and use for later are set here. 

```{r}
correct_RD <- function(data, output_path){
  
  ColumnNames <- c("sp","treeid","setTleaf",
             "anet.21p",    
             "anet.0p",
             "Rd",
             "corranet",
             "anet.delta",
             "pr.CO2", 
             "pr.real",
             "pr.percent",    
             "gsw.21p",    
             "gsw.0p",     
             "gsw.delta",  
             "gsw.percent",
             "E.21p",     
             "E.0p",      
             "E.delta",   
             "E.percent", 
             "ETR.21p",     
             "ETR.0p",
             "ETR.delta",   
             "ETR.percent",
             "JT",
             "JO1",
             "JC1",
             "JO2",
             "JC2",
             "JO1.percent",
             "JC1.percent",
             "JO2.percent",
             "JC2.percent",
             "NPQ.21p",
             "NPQ.0p",
             "NPQ.delta",
             "NPQ.percent")
  
  output_data <- data.frame(matrix(nrow = 0, ncol = length(ColumnNames)))
  names(output_data) <- ColumnNames

  for(dataIdx in 1:length(data)){
    
    speciesdata <- data[[dataIdx]]
    if(nrow(speciesdata) != 2){
      print(names(dflist)[dataIdx])
      print(nrow(speciesdata))
      next
    }
    
    
    sp               <-     speciesdata$sp[1]
    treeid           <-     speciesdata$treeid[1]
    setTleaf         <-     speciesdata$setTleaf[1]
    
    p21              <-     subset(speciesdata, olev == "21p")    
    p0               <-     subset(speciesdata, olev == "0p")
    
    anet.21p         <-     p21$A
    anet.0p          <-     p0$A
     
    
    # Calculate corrected Anet. The values for slope and intercept       where calculated using values taken from the literature. 
    
    Rd               <-      0.05554*anet.21p+ 0.11395 
    corranet         <-      anet.21p-Rd          
    # "corranet" is Anet corrected for Rd
    
    
    anet.delta       <-      anet.0p - anet.21p
    pr.CO2           <-      anet.delta * 0.5
    pr.real          <-      anet.delta + pr.CO2
    pr.percent       <-      pr.real/anet.21p
    
    gsw.21p          <-      p21$gsw
    gsw.0p           <-      p0$gsw
    gsw.delta        <-      p0$gsw - p21$gsw
    gsw.percent      <-      gsw.delta/p21$gsw
    
    
    E.21p            <-      p21$E
    E.0p             <-      p0$E
    E.delta          <-      p0$E - p21$E
    E.percent        <-      E.delta/p21$E
    
    
    ETR.21p          <-      p21$ETR
    ETR.0p           <-      p0$ETR
    ETR.delta        <-      p0$ETR - p21$ETR
    ETR.percent      <-      ETR.delta/p21$ETR
    JT               <-      ETR.21p + ETR.0p
    JO1              <-      pr.percent * JT
    JC1              <-      JT-JO1
    JO2              <-      ((2/3)*(JT-4*(anet.21p)))
    JC2              <-      ((1/3)*(JT+8*(anet.21p)))
    JO1.percent      <-      JO1/JT
    JC1.percent      <-      JC1/JT
    JO2.percent      <-      JO2/JT
    JC2.percent      <-      JC2/JT
    
    
    NPQ.21p          <-      p21$NPQ
    NPQ.0p           <-      p0$NPQ
    NPQ.delta        <-      NPQ.0p - NPQ.21p
    NPQ.percent      <-      NPQ.delta/NPQ.21p
    
    new_data = data.frame(as.character(sp),
                          as.character(treeid),
                          as.character(setTleaf), 
                          anet.21p,    
                          anet.0p,     
                          Rd, 
                          corranet,
                          anet.delta,
                          pr.CO2,
                          pr.real,
                          pr.percent,
                          gsw.21p,    
                          gsw.0p,     
                          gsw.delta,  
                          gsw.percent,
                          p21$E,     
                          E.0p,      
                          E.delta,   
                          E.percent, 
                          ETR.21p,     
                          ETR.0p,   
                          ETR.delta,   
                          ETR.percent,
                          JT,
                          JO1,
                          JC1,
                          JO2,
                          JC2,
                          JO1.percent,
                          JC1.percent,
                          JO2.percent,
                          JC2.percent,
                          NPQ.21p,
                          NPQ.0p,
                          NPQ.delta, 
                          NPQ.percent)
    
    names(new_data) <- names(output_data)
    output_data <- rbind(output_data, new_data)
    
    
  }
  
  file_path = paste(output_path, "species-output.csv", sep = "/")
  if(!dir.exists(output_path)){
    dir.create(output_path, recursive = T)
  }
  
  
  index <- 1
  while(file.exists(file_path)){
    file_path <- paste(output_path, paste("species-output", as.character(index), ".csv", sep =""), sep = "/")
    index = index + 1
  }
 
  

  write.table(output_data, file = file_path, 
              row.names = FALSE, col.names = T, sep = ";")
}


correct_RD(dflist, "./")

```



#Loading in the output file of the for loop into the "outs" dataframe. 

```{r}

outs <- read.csv("species-output.csv", stringsAsFactors = T, sep = ";")

```


# Thereafter, we can make some figures with our compiled data, for instance plotting the absolute rates of photorespiration over the leaf temperature for each species. 

```{r}
library(dplyr)
s.table <- outs %>%
  group_by(sp, setTleaf) %>%
  summarise(pr.real.avg = mean(pr.real),
            se = sd(pr.real))


svg(filename = "pr-raw.svg", width = 16, height = 4.5, bg = "transparent")

ggplot(s.table, aes(y = pr.real.avg, x = setTleaf, group=sp))+
  geom_point(col="black", size = 1.5)+
  geom_errorbar(aes(ymin=pr.real.avg-se, ymax=pr.real.avg+se), width=.2)+ #pmin can be used to cap the ymax. (e.g., pmin(pr.real.avg+se, 1.0))
  geom_smooth(data = outs, mapping = aes(y=pr.real, x=setTleaf), se = F, method="lm")+
  ggpmisc::stat_poly_eq(
    data = outs,
    formula = y ~ x,
    aes(x = setTleaf, y = pr.real, label = paste(after_stat(p.value.label), sep = "")),
    label.x = 20,
    label.y = 1.5,
    digits = 2
  )+
  # stat_regline_equation(data = outs, aes(x = setTleaf, y=pr.real, label = paste(..eq.label.., ..adj.rr.label.., paste("p = ", ..p.value..), sep = "~~~")),
  # formula = y~x,
  # label.x =22, label.y = 1.5)+
  facet_wrap(~sp, ncol = 7) +
  scale_y_continuous(limits = c(0, 15), name = "Net CO2 assimilation rate (0% O2) - Net CO2 assimilation rate (21 % O2)")+
  scale_x_continuous(limits = c(20,40), name = "Leaf Temperature (°C)")+
  ggthemes::theme_base()
dev.off()
```

# Here we are running a for loop testing the relationship between species, setTleaf and the combined effect of them (I.e the temperature response of species), against pr.real (Absolute photorespiration rates ("10")), pr.percent (Proportional photorespiration rates ("11")), anet.21p (Ambient photosynthesis rates ("4")), anet.0p (O2-free photosynthesis rates ("5")) and JO2.percent (proportional oxygenation rates calculated using ETR only ("31)).

```{r}

colindeces <- c(10, 11, 4, 5, 31)
method.slope <- c()
value.slope <- c()
species.slope <- c()

names(outs)[colindeces[1]]

for (index in 1:length(colindeces)){
  
  
  independentVariable = outs[,colindeces[index]] 
  
  mod1 <- nlme::lme(independentVariable ~  sp , data = outs, 
                    random = ~1|treeid, 
                    method = "REML", 
                    na.action=na.omit) ; anova(mod1)
  print(paste("This analysis is based on :", names(outs)[colindeces[index]]))
  print(summary(mod1))
  
  mod1 <- nlme::lme(independentVariable ~  setTleaf , data = outs, 
                    random = ~1|sp, 
                    method = "REML", 
                    na.action=na.omit) ; anova(mod1)
  
  print(paste("This analysis is based on :", names(outs)[colindeces[index]]))
  print(summary(mod1))
  mod1 <- nlme::lme(independentVariable ~  sp * setTleaf, 
                    data = outs, 
                    random = ~1|treeid, 
                    method = "REML", 
                    na.action=na.omit) ; anova(mod1)
  
  print(paste("This analysis is based on :", names(outs)[colindeces[index]]))
  print(summary(mod1))
  
  
  print("----------------------------------------------------------------")
}


library(tidyr)

long_data <- outs %>%
  gather(key = "method", value = "y", pr.percent, JO2.percent)

# Fit a model with an interaction term between method and setTleaf
long_model <- lm(y ~ sp * setTleaf * method, data = long_data)

# Check the model summary
summary(long_model)


# Fit the model with and without the interaction term between method and setTleaf
model_interaction <- lm(y ~ sp * setTleaf * method, data = long_data)
model_no_interaction <- lm(y ~ sp * setTleaf + method, data = long_data)

# Compare the models using a likelihood ratio test
anova(model_no_interaction, model_interaction)

# Visualize the comparison
library(ggplot2)
ggplot(long_data, aes(x = setTleaf, y = y, color = method)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression lines
  facet_wrap(~sp)+
  labs(title = "Comparison of Slopes for y1 and y2") +
  theme_minimal()

```



# Post hoc test ------- HSD, honest significant difference, I still need to figure out how 

```{r}

m7 <- aov(anet.0p ~ sp, data = outs, random = ~1|treeid, method = "REML", na.action=na.omit) ; anova(m7)
par(las=2)
par(mar=c(8,8,2,2)) 
plot(TukeyHSD(m7))


t1 <- TukeyHSD(m7)
t1 <- as.data.frame(t1$sp)
t1$com <- rownames(t1)

rownames(t1) <- NULL
colnames(t1) <- c("prdiff","lwr", "upr","padj", "com")
blah <- subset(t1, padj <= 0.05)
```

```{r}

```

```{r}

```

