---
title: "Photorespiration-Species"
author: "Philip"
date: "2024-11-06"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# 1. Asign all the variable labels that will be useful for plotting the data.

```{r}
ytleaf          <- expression("T"["leaf"]~degree~"C")
photorate       <- expression("" ~mu ~ mol ~ "CO"[2]~m^{-2} ~ "s"^{-1})
labrsq          <- expression(R^{2}~"=")
labvpdl         <- "VPD (kPa)"
lab_photo       <- expression(paste(italic(A)[Net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')'))
lab_photo1      <- expression(paste(italic(A)[Net], ' (', mu * 'mol m'^{-2}*'s'^{-1}*')'))
lab_Tleaf       <- expression(paste("Leaf temperature (", degree,"C)"))
labcond         <- expression("g"["s"]~~"in"~mol~H[2]*O~m^{-2}~s^{-1})
labcond1        <- expression("Conductance to "~H[2]*O~"in"~mol~H[2]*O~m^{-2}~s^{-1})
labtrmmol       <- expression("E"~~"in"~mol~H[2]*O~m^{-2}~s^{-1})
etrrate         <- expression("" ~mu ~ mol ~"photon" ~m^{-2} ~ "s"^{-1})
photorate       <- expression("" ~mu ~ mol ~ "CO"[2]~m^{-2} ~ "s"^{-1})
labgs           <- expression(paste(italic(g[s])*" ("~mol[H[2]*O]~m^{-2}~s^{-1}~")"))
```


# 2. Loading in the compiled species data into a dataframe (df). If the output file is already created, go directly to entry 5.

```{r}
df <- read.csv("Data/Uppsala-2024-Summer-Photorespiration-SpotMes-TreeSpp.csv", header = T, stringsAsFactors = T, sep = ";")
```



#3. Clean up of the dataframe by sorting out any entries in the loaded file that will not be used in the analysis. Thereafter pairing entries together (O2-free with ambient for the same replicate). Then converting all the pairs into factors.

```{r}
df     <- subset(df, dq == "yes") # eliminates unreliable values.

df$pairid <- paste0(df$sprp, "-", df$setTleaf)

df$pairid <- as.factor(df$pairid)


length(unique(df$pairid))       # individual runs + season

dflist    <-   split(df,list(df$pairid))
dflist    <-   split(df,unique(list(df$pairid)))
dflist    <-   dflist[sapply(dflist, nrow)>0] 

setwd("output")
```



#4. The for loop that will give us our worksheet for the statistical analysis. All data that we want to calculate and use for later are made here. 

```{r}
correct_RD <- function(data, output_path){
  
  ColumnNames <- c("sp","treeid","setTleaf",
             "anet.21p",    
             "anet.0p",
             "Rd",
             "corranet",
             "anet.delta",
             "pr.CO2", 
             "pr.real",
             "pr.percent",    
             "gsw.21p",    
             "gsw.0p",     
             "gsw.delta",  
             "gsw.percent",
             "E.21p",     
             "E.0p",      
             "E.delta",   
             "E.percent", 
             "ETR.21p",     
             "ETR.0p",
             "ETR.delta",   
             "ETR.percent",
             "JT",
             "JO1",
             "JC1",
             "JO2",
             "JC2",
             "JO1.percent",
             "JC1.percent",
             "JO2.percent",
             "JC2.percent",
             "NPQ.21p",
             "NPQ.0p",
             "NPQ.delta",
             "NPQ.percent")
  
  output_data <- data.frame(matrix(nrow = 0, ncol = length(ColumnNames)))
  names(output_data) <- ColumnNames

  for(dataIdx in 1:length(data)){
    
    speciesdata <- data[[dataIdx]]
    if(nrow(speciesdata) != 2){
      print(names(dflist)[dataIdx])
      print(nrow(speciesdata))
      next
    }
    
    
    sp               <-     speciesdata$sp[1]
    treeid           <-     speciesdata$treeid[1]
    setTleaf         <-     speciesdata$setTleaf[1]
    
    p21              <-     subset(speciesdata, olev == "21p")    
    p0               <-     subset(speciesdata, olev == "0p")
    
    anet.21p         <-     p21$A
    anet.0p          <-     p0$A
     
    
    # Calculate corrected Anet. The values for slope and intercept       where calculated using values taken from the literature. 
    
    Rd               <-      0.05554*anet.21p+ 0.11395 
    corranet         <-      anet.21p-Rd          
    # "corranet" is Anet corrected for Rd
    
    
    anet.delta       <-      anet.0p - anet.21p
    pr.CO2           <-      anet.delta * 0.5
    pr.real          <-      anet.delta + pr.CO2
    pr.percent       <-      pr.real/anet.21p
    
    gsw.21p          <-      p21$gsw
    gsw.0p           <-      p0$gsw
    gsw.delta        <-      p0$gsw - p21$gsw
    gsw.percent      <-      gsw.delta/p21$gsw
    
    
    E.21p            <-      p21$E
    E.0p             <-      p0$E
    E.delta          <-      p0$E - p21$E
    E.percent        <-      E.delta/p21$E
    
    
    ETR.21p          <-      p21$ETR
    ETR.0p           <-      p0$ETR
    ETR.delta        <-      p0$ETR - p21$ETR
    ETR.percent      <-      ETR.delta/p21$ETR
    JT               <-      ETR.21p + ETR.0p
    JO1              <-      pr.percent * JT
    JC1              <-      JT-JO1
    JO2              <-      ((2/3)*(JT-4*(anet.21p)))
    JC2              <-      ((1/3)*(JT+8*(anet.21p)))
    JO1.percent      <-      JO1/JT
    JC1.percent      <-      JC1/JT
    JO2.percent      <-      JO2/JT
    JC2.percent      <-      JC2/JT
    
    
    NPQ.21p          <-      p21$NPQ
    NPQ.0p           <-      p0$NPQ
    NPQ.delta        <-      NPQ.0p - NPQ.21p
    NPQ.percent      <-      NPQ.delta/NPQ.21p
    
    new_data = data.frame(as.character(sp),
                          as.character(treeid),
                          as.character(setTleaf), 
                          anet.21p,    
                          anet.0p,     
                          Rd, 
                          corranet,
                          anet.delta,
                          pr.CO2,
                          pr.real,
                          pr.percent,
                          gsw.21p,    
                          gsw.0p,     
                          gsw.delta,  
                          gsw.percent,
                          p21$E,     
                          E.0p,      
                          E.delta,   
                          E.percent, 
                          ETR.21p,     
                          ETR.0p,   
                          ETR.delta,   
                          ETR.percent,
                          JT,
                          JO1,
                          JC1,
                          JO2,
                          JC2,
                          JO1.percent,
                          JC1.percent,
                          JO2.percent,
                          JC2.percent,
                          NPQ.21p,
                          NPQ.0p,
                          NPQ.delta, 
                          NPQ.percent)
    
    names(new_data) <- names(output_data)
    output_data <- rbind(output_data, new_data)
    
    
  }
  
  file_path = paste(output_path, "species-output.csv", sep = "/")
  if(!dir.exists(output_path)){
    dir.create(output_path, recursive = T)
  }
  
  
  index <- 1
  while(file.exists(file_path)){
    file_path <- paste(output_path, paste("species-output", as.character(index), ".csv", sep =""), sep = "/")
    index = index + 1
  }
 
  

  write.table(output_data, file = file_path, 
              row.names = FALSE, col.names = T, sep = ";")
}


correct_RD(dflist, "./")

```



#5. Loading in the output file of the for loop into the "outs" dataframe. 

```{r}

outs <- read.csv("species-output.csv", stringsAsFactors = T, sep = ";")

```


#6. Before plotting any figures, it these packages need to be loaded in. "dplyr" might need to be loaded in a separate time after all of them have been loaded once to actually work. 

```{r}
library(stats)
library(base)
library(dplyr)
library(ggplot2)
```



###Figures



#7. Absolute photorespiration rates across temperature
```{r}
svg(filename = "pr-raw.svg", width = 16, height = 4.5, bg = "transparent")


absolute.table <- outs %>%
  group_by(sp, setTleaf) %>%
  summarise(pr.real.avg = mean(pr.real),
            se = sd(pr.real))

ggplot(absolute.table, aes(y = pr.real.avg, x=setTleaf, group=sp))+
  geom_errorbar(aes(ymin=pr.real.avg-se, ymax=pr.real.avg+se),col ="grey70", width= 2.5)+ #pmin can be used to cap the ymax. (e.g., pmin(pr.real.avg+se, 1.0))
  geom_smooth(data = outs, mapping = aes(y=pr.real, x=setTleaf), se = F, method="lm", col = "grey80")+
  geom_point(col="grey30", size = 3, pch = 21, fill = "cornflowerblue", stroke = 0.85)+
  facet_wrap(~sp, ncol = 7) +
  scale_y_continuous(limits = c(0, 15), name = expression(paste(italic(R)[p], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  scale_x_continuous(limits = c(20,40), name = "Leaf Temperature (°C)")+
  ggthemes::theme_base() +
theme(axis.text.y = element_text(size = 15), 
      axis.text.x = element_text(size = 15),
      panel.border = element_rect(color = "grey70"))
dev.off()

```



#8. Absolute photosynthesis across temperature

```{r}
svg(filename = "anet21-raw.svg", width = 16, height = 4.5, bg = "transparent")


anet.table <- outs %>%
  group_by(sp, setTleaf) %>%
  summarise(anet.21p.avg = mean(anet.21p),
            se = sd(anet.21p))


ggplot(anet.table, aes(y = anet.21p.avg, x=setTleaf, group=sp))+
   #pmin can be used to cap the ymax. (e.g., pmin(anet.21p.avg+se, 1.0))
  geom_smooth(data = outs, mapping = aes(y=anet.21p, x=setTleaf), se = F, method="lm", col = "grey80")+
  geom_errorbar(aes(ymin=anet.21p.avg-se, ymax=anet.21p.avg+se),col ="grey70", width= 2.5) +
  geom_point(col="grey30", size = 3, pch = 21, fill = "limegreen", stroke = 0.85)+
  # stat_regline_equation(data = outs, aes(x = setTleaf, y=anet.21p, label = paste(..eq.label.., ..adj.rr.label.., paste("p = ", ..p.value..), sep = "~~~")),
  # formula = y~x,
  # label.x =22, label.y = 1.5)+
  facet_wrap(~sp, ncol = 7) +
  scale_y_continuous(limits = c(0, 25), name = expression(paste(italic(A)[Net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  scale_x_continuous(limits = c(20,40), name = "Leaf Temperature (°C)")+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15),
        panel.border = element_rect(color = "grey70"))
dev.off()

```


#9. Percentage of photorespiration across temperature

```{r}
svg(filename = "pr-percent.svg", width = 16, height = 4.5, bg = "transparent")



percent.table <- outs %>%
  group_by(sp, setTleaf) %>%
  summarise(pr.percent.avg = mean(pr.percent),
            se = sd(pr.percent))


ggplot(percent.table, aes(y = pr.percent.avg, x=setTleaf, group=sp))+
  #pmin can be used to cap the ymax. (e.g., pmin(pr.real.avg+se, 1.0))
  geom_smooth(data = outs, mapping = aes(y=pr.percent, x=setTleaf), se = F, method="lm", col = "grey80")+
  # ggpmisc::stat_poly_eq(
  #   data = outs,
  #   formula = y ~ x,
  #   aes(x = setTleaf, y = pr.percent, label = paste(after_stat(p.value.label), sep = "~~~")),
  #   label.x = 22,
  #   label.y = 1.5,
  # )+
  geom_errorbar(aes(ymin=pr.percent.avg-se, ymax=pr.percent.avg+se),col ="grey70", width= 2.5)+
  geom_point(col="grey30", size = 3, pch = 21, fill = "red3", stroke = 0.85)+
  # stat_regline_equation(data = outs, aes(x = setTleaf, y=pr.real, label = paste(..eq.label.., ..adj.rr.label.., paste("p = ", ..p.value..), sep = "~~~")),
  # formula = y~x,
  # label.x =22, label.y = 1.5)+
  facet_wrap(~sp, ncol = 7)+
  scale_y_continuous(limits = c(0, 1.5), name = 
                       expression(paste(italic(R)[p]/italic(A)[Net])))+
  scale_x_continuous(limits = c(20,40), name = "Leaf Temperature (°C)")+
  ggthemes::theme_base() +
  theme(axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15),
        panel.border = element_rect(color = "grey70"))
dev.off()

```


#10. Creating a factorial dataframe of our leaf temperatures used for plotting

```{r}
outs$tl <- as.factor(outs$setTleaf)
```


#11. Absolute photorespiration and photosynthesis rates across species boxplot (across all temperatures)

```{r}
svg(filename = "Rp-species.svg", width = 12, height = 6, bg = "transparent")

ggplot(outs, aes(y = pr.real, x=sp)) +
  geom_boxplot(aes(x = sp, y = pr.real), outliers =  F)  +
  geom_jitter(aes(size = tl),  color = "grey", alpha = 0) +
  
  xlab(" ") +
  scale_y_continuous(limits = c(0,15), 
                     name = expression(paste(italic(R)[p], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15),
        panel.border = element_rect(color = "grey70")) +
  coord_flip()
dev.off()

svg(filename = "Anet-species.svg", width = 12, height = 6, bg = "transparent")

ggplot(outs, aes(y = anet.21p, x=sp)) +
  geom_boxplot(aes(x = sp, y = anet.21p), outliers =  F)  +
  geom_jitter(aes(size = tl),  color = "grey", alpha = 0) +
  
  xlab(" ") +
  scale_y_continuous(limits = c(0,25), 
                     name = expression(paste(italic(A)[net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15),
        panel.border = element_rect(color = "grey70")) +
  coord_flip()
dev.off()



```


#12. Absolute photorespiration and photosynthesis rates across species including the spread of temperature points 

```{r}
svg(filename = "Rp-species-spotted.svg", width = 12, height = 6, bg = "transparent")

ggplot(outs, aes(y = pr.real, x=sp)) +
  geom_boxplot(aes(x = sp, y = pr.real), outliers =  F)  +
  geom_jitter(aes(size = tl), color = as.factor(outs$setTleaf), alpha = 0.5) +
  xlab(" ") +
  scale_y_continuous(limits = c(0,15), 
                     name = expression(paste(italic(R)[p], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 20), 
        axis.text.x = element_text(size = 20),
        panel.border = element_rect(color = "grey70")) +
  coord_flip()
dev.off()


svg(filename = "Anet-species-spotted.svg", width = 12, height = 6, bg = "transparent")

ggplot(outs, aes(y = anet.21p, x=sp)) +
  geom_boxplot(aes(x = sp, y = anet.21p), outliers =  F)  +
  geom_jitter(aes(size = tl), color = as.factor(outs$setTleaf), alpha = 0.5) +
  xlab(" ") +
  scale_y_continuous(limits = c(0,25), 
                     name = expression(paste(italic(A)[net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 20), 
        axis.text.x = element_text(size = 20),
        panel.border = element_rect(color = "grey70")) +
  coord_flip()
dev.off()
```



#13. Correlation between photorespiration and photosynthesis rates

```{r}
svg(filename = "Rp-over-Anet.svg", width = 7, height = 4.5, bg = "transparent")


ggplot(outs, aes(y = pr.real, x=anet.21p)) +
  geom_point(size = 3, 
             stroke = 0.85, aes(color = sp, shape = sp)) +
  geom_smooth(se = T, method="lm", col = "grey50") +
  scale_x_continuous(limits = c(0, 25), 
                     name = expression(paste(italic(A)[Net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')'))) +
  scale_y_continuous(limits = c(0,25), 
                     name = expression(paste(italic(R)[p], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15),
        panel.border = element_rect(color = "grey70")) +
  scale_shape_manual(values = c(21,22,23,24,25,1,2)) +
  geom_abline(slope = 1, linetype = "dashed", color = "grey50")
dev.off()

```

#14. Correlation between photorespiration and photosynthesis rates for each leaf temperature point

```{r}
svg(filename = "Rp-over-Anet-alltemp.svg", width = 16, height = 4.5, bg = "transparent")


ggplot(outs, aes(y = pr.real, x=anet.21p, color = sp)) +
  geom_smooth(data = outs, mapping = aes(y=pr.real, x=anet.21p), 
              se = T, method="lm", col = "grey80") +
  geom_point(size = 3, 
             stroke = 0.85, aes(color = sp, shape = sp)) +
  scale_x_continuous(limits = c(0, 25), 
                     name = expression(paste(italic(A)[Net], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')'))) +
  scale_y_continuous(limits = c(0,25), 
                     name = expression(paste(italic(R)[p], ' (', mu * ~'mol'~ " CO"[2]~' m'^{-2}*' s'^{-1}*')')))+
  ggthemes::theme_base()+
  theme(axis.text.y = element_text(size = 15), 
        axis.text.x = element_text(size = 15),
        panel.border = element_rect(color = "grey70")) +
  scale_shape_manual(values = c(21,22,23,24,25,1,2))+
  geom_abline(slope = 1, linetype = "dashed", color = "grey50") +
  facet_wrap(~setTleaf)
dev.off()
```



### Statistical tests



#15. testing absolute photorespiration rates correlation to: Rp1: species, Rp2: temperature and Rp3: species * temperature interaction. 
```{r}
Rp1 <- nlme::lme(pr.real ~  sp, data = outs, 
                  random = ~1|treeid, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Rp1)

Rp2 <- nlme::lme(pr.real ~  setTleaf, data = outs, 
                  random = ~1|sp, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Rp2)

Rp3 <- nlme::lme(pr.real ~  sp * setTleaf, data = outs, 
                  random = ~1|treeid, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Rp3)
```

#16. testing absolute photosynthesis rates correlation to: Anet1: species, Anet2: temperature and Anet3: species * temperature interaction. 

```{r}
Anet1 <- nlme::lme(anet.21p ~  sp, data = outs, 
                  random = ~1|treeid, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Anet1)

Anet2 <- nlme::lme(anet.21p ~  setTleaf, data = outs, 
                  random = ~1|sp, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Anet2)

Anet3 <- nlme::lme(anet.21p ~  sp * setTleaf, data = outs, 
                  random = ~1|treeid, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Anet3)
```

#17. testing percentage photorespiration rates correlation to: PR1: species, PR2: temperature and PR3: species * temperature interaction. 

```{r}
PR1 <- nlme::lme(pr.percent ~  sp, data = outs, 
                  random = ~1|treeid, 
                  method = "REML", 
                  na.action=na.omit) ; anova(PR1)

PR2 <- nlme::lme(pr.percent ~  setTleaf, data = outs, 
                  random = ~1|sp, 
                  method = "REML", 
                  na.action=na.omit) ; anova(PR2)


PR3 <- nlme::lme(pr.percent ~  sp * setTleaf, data = outs, 
                  random = ~1|treeid, 
                  method = "REML", 
                  na.action=na.omit) ; anova(PR3)

```


#18. Creating dataframes for each temperature
```{r}
outs25 <- subset(outs, setTleaf == 25)
outs30 <- subset(outs, setTleaf == 30)
outs35 <- subset(outs, setTleaf == 35)
```

#19. Relation between photorespiration and photosynthesis for all and split each temperature point

```{r}
Rp-Anet <- nlme::lme(pr.real ~  anet.21p, data = outs, 
                 random = ~1|sp, 
                 method = "REML", 
                 na.action=na.omit) ; anova(Rp-Anet)


Rp-Anet25 <- nlme::lme(pr.real ~  anet.21p, data = outs25, 
                  random = ~1|sp, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Rp-Anet25)

Rp-Anet30 <- nlme::lme(pr.real ~  anet.21p, data = outs30, 
                  random = ~1|sp, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Rp-Anet30)

Rp-Anet <- nlme::lme(pr.real ~  anet.21p, data = outs35, 
                  random = ~1|sp, 
                  method = "REML", 
                  na.action=na.omit) ; anova(Rp-Anet35)
```





### WIP



# Post hoc test ------- HSD, honest significant difference, I still need to figure out how 

```{r}

m7 <- aov(anet.0p ~ sp, data = outs, random = ~1|treeid, method = "REML", na.action=na.omit) ; anova(m7)
par(las=2)
par(mar=c(8,8,2,2)) 
plot(TukeyHSD(m7))


t1 <- TukeyHSD(m7)
t1 <- as.data.frame(t1$sp)
t1$com <- rownames(t1)

rownames(t1) <- NULL
colnames(t1) <- c("prdiff","lwr", "upr","padj", "com")
blah <- subset(t1, padj <= 0.05)
```

```{r}

```

```{r}

```

